BIOL 274 Homework 2

### 2024-02-01 bio_dynamics/bio274_hw2.Rmd
hw2 for bio, hw1 for math

---
title: "BIOL 274 Homework 2"
author: "Ruby Krasnow"
date: "2024-01-31"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

@@

Let's see what we got!
```{r,   fig.height=3, fig.width=5, fig.width=5}
ggplot()+
  geom_line(data=fish_output, aes(x=time, y=P, color=names_full))+
  theme_bw()+
  labs(x="Time (years)", y="Pop. size", color=NULL)
```

@@

ggplot()+
  geom_line(aes(x=t, y=y, color="P(t)"))+
  theme_bw()+
  labs(x="time", y="P")
```

@@

ggplot()+
  geom_line(aes(x=y, y=eval(dy), color="calculated"), linewidth=3, alpha=0.5)+
  geom_line(aes(x=y, y= 0.3*y*(1-(y/2500))-y/3, color="original dP/dt"))+
  scale_color_manual(values=c("lightblue", "black"))+
  theme_bw()+
  labs(x="P", y="dP/dt")
```

## Question 2 - for practice
Using Bernoulli's method, find a solution to the IVP representing the situation in #1b.

@@

ggplot()+
  geom_hline( linetype="dashed", aes(color="N0/a",yintercept=N0/a))+
  geom_line(aes(x=x, y=(N0*p0_small)/(a*p0_small+(N0-a*p0_small)*exp(-N0*x)), color="p0 << N0/a"))+
  geom_line(aes(x=x, y=(N0*p0_big)/(a*p0_big+(N0-a*p0_big)*exp(-N0*x)), color="p0 > N0/a"))+
  theme_bw()+
  labs(x="t", y="P", color=NULL)+
  scale_color_manual(values=c("gray", "#0f85a0", "#dd4124"))+
  ylim(0, NA)
```

@@

@@

```{r,   fig.height=3, fig.width=5  }
ggplot(data = euler_output)+
  geom_point(aes(x=x, y=y, color=as.factor(h)))+
  geom_line(aes(x=x, y=y, color=as.factor(h)), alpha=0.5)+
  theme_light()+
  labs(x="t", y="w", color="h")
```

@@

ggplot()+
  geom_line(data=actual_eqn, aes(x=t, y=w, color="Actual solution"), linewidth=1.5)+
  geom_point(data = euler_output, aes(x=x, y=y, color=as.factor(h)), size=1)+
  geom_line(data = euler_output, aes(x=x, y=y, color=as.factor(h)), alpha=0.5, linewidth=1)+
  theme_light()+
  labs(x="t", y="w", color="h")+
  scale_color_manual(values=c( "#e69b99","#89689d","#015b58", "black"))
#Credit: hex codes are from Jake Lawlor's R package PNWcolors
```

@@

```{r}
library(PNWColors)

euler_output_2<- c(1, 0.5, 0.1, 0.05) %>% 
  map(\(x) euler_fun(new_f, x0=0, y0=2, h=x, N=2)) %>% 
  bind_rows()

ggplot()+
  geom_point(data = euler_output_2, aes(x=x, y=y, color=as.factor(h)))+
  geom_line(data = euler_output_2, aes(x=x, y=y, color=as.factor(h)), alpha=0.8)+
  theme_light()+
  labs(x="t", y="y", color="h")+
  scale_color_manual(values=pnw_palette(name="Starfish",n=4,type="discrete"))
```

>>>>

---
title: "BIOL 274 Homework 2"
author: "Ruby Krasnow"
date: "2024-02-01"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(PNWColors)
```

@@

Let's see what we got!
```{r,   fig.height=3, fig.width=5, fig.width=5}
ggplot()+
  geom_line(data=fish_output, aes(x=time, y=P, color=names_full))+
  theme_bw()+
  labs(x="Time (years)", y="Pop. size", color=NULL)+
  scale_color_manual(values=pnw_palette(name="Sailboat",n=4,type="discrete"))
```

@@

ggplot()+
  geom_line(aes(x=t, y=y, color="P(t)"))+
  theme_bw()+
  labs(x="time", y="P", color=NULL)
```

@@

ggplot()+
  geom_line(aes(x=y, y=eval(dy), color="Calculated derivative"), linewidth=3, alpha=0.5)+
  geom_line(aes(x=y, y= 0.3*y*(1-(y/2500))-y/3, color="Original dP/dt"))+
  scale_color_manual(values=c("lightblue", "black"))+
  theme_bw()+
  labs(x="P", y="dP/dt", color=NULL)
```

## Question 2 again - for practice with Bernoulli's method
Using Bernoulli's method, find a solution to the IVP representing the situation in #1b.

@@

ggplot()+
  geom_hline( linetype="dashed", aes(color="N0/a",yintercept=N0/a))+
  geom_line(aes(x=x, y=(N0*p0_small)/(a*p0_small+(N0-a*p0_small)*exp(-N0*x)), color="p0 << N0/a"))+
  geom_line(aes(x=x, y=(N0*p0_big)/(a*p0_big+(N0-a*p0_big)*exp(-N0*x)), color="p0 > N0/a"))+
  theme_bw()+
  labs(x="t", y="P", color=NULL)+
  scale_color_manual(values=c("gray",pnw_palette(name="Bay",n=2,type="discrete")))+
  ylim(0, NA)
```

@@

The equations needed to use Euler's method are:
$$t_{n+1}=t_n+h$$
$$y_{n+1}=y_n+f(t_n, y_n)h$$
where $f(t_n, y_n)$ is the slope we calculate by plugging $t_n, y_n$ into our original DE.

@@

```{r,   fig.height=3, fig.width=5  }
ggplot(data = euler_output)+
  geom_point(aes(x=x, y=y, color=as.factor(h)))+
  geom_line(aes(x=x, y=y, color=as.factor(h)), alpha=0.5)+
  theme_light()+
  labs(x="t", y="w", color="h")+
  scale_color_manual(values=pnw_palette(name="Sailboat",n=3,type="discrete"))
```

@@

ggplot()+
  geom_line(data=actual_eqn, aes(x=t, y=w, color="Actual solution"), linewidth=1)+
  geom_point(data = euler_output, aes(x=x, y=y, color=as.factor(h)), size=1)+
  geom_line(data = euler_output, aes(x=x, y=y, color=as.factor(h)), alpha=0.5, linewidth=1)+
  theme_light()+
  labs(x="t", y="w", color="h")+
  scale_color_manual(values=c(pnw_palette(name="Sailboat",n=3,type="discrete"), "black"))
```

@@

```{r}
euler_output_2<- c(1, 0.5, 0.1, 0.05) %>% 
  map(\(x) euler_fun(new_f, x0=0, y0=2, h=x, N=2)) %>% 
  bind_rows()

ggplot()+
  geom_point(data = euler_output_2, aes(x=x, y=y, color=as.factor(h)))+
  geom_line(data = euler_output_2, aes(x=x, y=y, color=as.factor(h)), alpha=0.8)+
  theme_light()+
  labs(x="t", y="y", color="h")+
  scale_color_manual(values=pnw_palette(name="Sailboat",n=4,type="discrete"))
```
